---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# imputeGeneric

<!-- badges: start -->
[![R-CMD-check](https://github.com/torockel/imputeGeneric/workflows/R-CMD-check/badge.svg)](https://github.com/torockel/imputeGeneric/actions)
<!-- badges: end -->

The goal of imputeGeneric is to ease the implementation of imputation functions.

## Installation

You can install the development version of imputeGeneric from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("torockel/imputeGeneric")
```

## Purpose
The aim of `imputeGeneric` is to make the implementation and usage of imputation
methods easier. The main function of the package is `impute_iterative()`. With
this function close to any `parsnip` model can be turned into an imputation
method. Furthermore, other approaches can be used in a general imputation
framework. For more information, see the documentations of `impute_iterative()`,
`impute_supervised()` and `impute_unsupervised()`.

## Examples

The use of a `parsnip` model for imputation is demonstrated using regression trees
from `rpart` via `parsnip` (`decision_tree("regression")`). First, a data set with missing values is created. Then, this data
set is imputed once with regression trees using only completely observed rows
and columns for the model building.

``` {r example1}
library(imputeGeneric)
library(parsnip)
# create data set
set.seed(123)
ds_mis <- data.frame(X = rnorm(100), Y = rnorm(100))
ds_mis$Z <- 5 + 2* ds_mis$X + ds_mis$Y + rnorm(100)
ds_mis$Z[sample.int(100, 30)] <- NA
ds_mis$Y[sample.int(100, 20)] <- NA
# impute data set
ds_imp <- impute_iterative(ds_mis, decision_tree("regression"), max_iter = 1)
anyNA(ds_imp)
```
To use other `parsnip` models instead of regression trees, only the
`model_spec_parsnip` must be altered. E. g. for linear regression instead of
regression trees use `linear_reg()`.

``` {r example_lm}
ds_imp_lm <- impute_iterative(ds_mis, linear_reg(), max_iter = 1)
anyNA(ds_imp_lm)
```

Many aspects of the imputation can be specified and customized.  

For a more complex example, the missing values can be initially imputed with per
column mean values (`initial_imputation_fun = missMethods::impute_mean`). In
addition, all objects and columns can be used for the imputation models
(`rows_used_for_imputation = "all"` and `cols_used_for_imputation = "all"`).
Furthermore, the imputation can be iterative. The iterations will be stopped, if
either the difference between two imputed data sets falls below a threshold
(`stop_fun = stop_ds_difference, stop_fun_args = list(eps = 0.1)`) or the
maximum number of iterations (`max_iter = 5`) is reached.

``` {r example2}
ds_imp2 <- impute_iterative(
  ds_mis, decision_tree("regression"), 
  initial_imputation_fun = missMethods::impute_mean,
  cols_used_for_imputation = "all",
  rows_used_for_imputation = "all",
  stop_fun = stop_ds_difference,
  stop_fun_args = list(eps = 0.1),
  max_iter = 5)
anyNA(ds_imp2)
```
